draftFiguresPaper

### 2025-10-16 draftFiguresPaper.Rmd
summmary dataset to easily draft plots

@@

>>>>

```{r putting data together}

datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")

datasetPollution_avg <- datasetPollution %>% 
  filter(Year == 2019) %>% 
  group_by(Region) %>% 
  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))

datasetPollution_ctr <- datasetPollution_avg %>%
  mutate(
    iso3 = countrycode(Region,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(!is.na(iso3)) %>% 
  rename(countryName = Region)

papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)

fullCountrylist <- unique(countrycode::codelist$iso3c)

papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)

iso3_un_members <- c(
  "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
  "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
  "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
  "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
  "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
  "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
  "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
  "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
  "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
  "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
  "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
  "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
  "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
  "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
  "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
  "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
  "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
  "VUT","VEN","VNM","YEM","ZMB","ZWE"
)

papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
  filter(!is.na(iso3))

## adding survey nb

# 2) get total surveys by country
surveys_by_country <- surveyCountries_long_clean %>%
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

papers_by_country_pollution_filteredUN_survey <- merge(papers_by_country_pollution_filteredUN, surveys_by_country, by = "iso3", all.x = T, all.y = T)

papers_by_country_pollution_filteredUN_survey$n_surveys <- ifelse(is.na(papers_by_country_pollution_filteredUN_survey$n_surveys), 0, papers_by_country_pollution_filteredUN_survey$n_surveys)

population <- read.csv("dataNew/GHS_pop.csv", sep = ";")

popCtr <- population %>%
  group_by(GADM_ISO) %>% 
  summarize(population2015 = sum(X2015))

summaryDataset <- merge(papers_by_country_pollution_filteredUN_survey, popCtr, by.x = "iso3", by.y = "GADM_ISO", all.x = T, all.y = F)

summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
                  labels = c("very low", "low", "medium", "high", "very high"), 
                  right = FALSE)

plot(summaryDataset %>% 
       select(Income.group, pm25Avg, n_papers, n_surveys, population2015, categoryPollution))

summaryDataset$isChina <- ifelse(summaryDataset$iso3 == "CHN", 1, 0)
summaryDataset$isIndia <- ifelse(summaryDataset$iso3 == "IND", 1, 0)

write.csv(summaryDataset, "output/datasetPlots.csv", row.names = F)

@@

datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")

datasetPollution_avg <- datasetPollution %>% 
  filter(Year == 2019) %>% 
  group_by(Region) %>% 
  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))

datasetPollution_ctr <- datasetPollution_avg %>%
  mutate(
    iso3 = countrycode(Region,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(!is.na(iso3))

papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)

fullCountrylist <- unique(countrycode::codelist$iso3c)

papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)

papers_by_country_pollution_filtered <- subset(papers_by_country_pollution, iso3 %in% fullCountrylist) %>% 
  filter(!is.na(iso3))

papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()

papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth()

papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

subset(papers_by_country_pollution_filtered, n_papers > 0) %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

papers_by_country_pollution_filtered %>% 
  filter(n_papers > 0) %>%
  filter(iso3 != "CHN") %>% 
  ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

papers_by_country_pollution_filtered %>% 
  # filter(n_papers > 0) %>%
  filter(iso3 != "CHN") %>% 
  ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

iso3_un_members <- c(
  "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
  "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
  "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
  "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
  "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
  "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
  "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
  "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
  "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
  "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
  "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
  "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
  "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
  "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
  "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
  "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
  "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
  "VUT","VEN","VNM","YEM","ZMB","ZWE"
)

papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
  filter(!is.na(iso3))

papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()

papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth()

papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = (n_papers)))+
  geom_point()+
  geom_smooth(method = "lm")

papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = (n_papers)))+
  geom_point()+
  geom_smooth(method = "lm")+
  ylim(0,40)

subset(papers_by_country_pollution_filteredUN, n_papers > 0) %>% ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

papers_by_country_pollution_filteredUN %>% 
  filter(n_papers > 0) %>%
  filter(iso3 != "CHN") %>% 
  ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

papers_by_country_pollution_filteredUN %>% 
  # filter(n_papers > 0) %>%
  filter(iso3 != "CHN") %>% 
  ggplot(aes(x = pm25Avg, y = n_papers))+
  geom_point()+
  geom_smooth(method = "lm")

papers_by_country_pollution_filteredUN %>% ggplot(aes(x = n_papers > 0, y = pm25Avg))+
  geom_boxplot()

papers_by_country_pollution_filteredUN$categoryPollution <- cut(papers_by_country_pollution_filteredUN$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
                  labels = c("very low", "low", "medium", "high", "very high"), 
                  right = FALSE)

papers_by_country_pollution_filteredUN %>% 
  # filter(iso3 != "CHN") %>%
  ggplot(aes(x = categoryPollution, y = log(n_papers)))+
  geom_boxplot()

papers_by_country_pollution_filteredUN_summary <- papers_by_country_pollution_filteredUN %>% 
  mutate(paperBin = ifelse(n_papers > 0, 1, 0)) %>% 
  group_by(categoryPollution) %>% 
  summarize(proportionPaper = sum(paperBin, na.rm=T)/n(),
            nbCountries = n())

papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = proportionPaper))+
  geom_boxplot()+
  ylim(0,1)

papers_by_country_pollution_filteredUN_summary <- papers_by_country_pollution_filteredUN %>% 
  group_by(categoryPollution) %>% 
  summarize(papersPerCountry = sum(n_papers, na.rm=T)/n(),
            nbCountries = n(),
            medianPapersPerCountry = median(n_papers, na.rm=T))

papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = medianPapersPerCountry))+
  geom_boxplot()

table(papers_by_country_pollution_filteredUN$categoryPollution)

papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = papersPerCountry))+
  geom_boxplot()

TEST <- papers_by_country_pollution_filteredUN %>% 
  filter(categoryPollution == "very high")

# ylim(0,1)
```
